---
title: "Protected area effectiveness occupancy analysis"
author: "Rob Cooke"
date: "11/04/2020"
output: html_notebook
---

## Set-up ##

Here we load the necessary packages

```{r echo = FALSE}
knitr::opts_chunk$set(echo = TRUE)

# # install JAGS in terminal
# sudo apt install jags

library(remotes) # remotes: package installation
library(sf) # sf: spatial manipulation
library(dplyr) # dplyr: data manipulation
library(tidyr) # tidyr: data manipulation
library(rjags) # rjags: occupancy models
library(R2jags) # R2jags: occupancy models
library(raster) # raster: spatial manipulation
library(HDInterval) # HDInterval: credible intervals
library(effsize) # effsize: effect sizes
library(betapart) # betapart: temporal beta diversity
library(adespatial) # adespatial: temporal beta diversity
library(ggplot2) # ggplot2: plotting
library(cowplot) # cowplot: plotting

# # sparta package from github
# withr::with_envvar(c(R_REMOTES_NO_ERRORS_FROM_WARNINGS = "true"),
# remotes::install_github("BiologicalRecordsCentre/sparta", force = TRUE))
# library(sparta) # sparta: occupancy models

# # folder saved locally to packrat
# library(BRCmap) # BRCmap: grid references

# # wrappeR package from github
# remotes::install_github("https://github.com/BiologicalRecordsCentre/wrappeR", ref = "main")
library(wrappeR) # wrappeR: multi-species indicators

# # install gghalves from github
# remotes::install_github('erocoar/gghalves')
library(gghalves)

# # install occAssess from github
# devtools::install_github("https://github.com/robboyd/occAssess")
library(occAssess)

# # install igraph from github
# # problems with CRAN version similar to this https://github.com/igraph/rigraph/issues/335
# remotes::install_github("igraph/rigraph@master")

# # commit packages to packrat library
# packrat::snapshot()
# packrat::status()

# set location of temporary directory for raster package
rasterOptions(tmpdir = "temp")

# source pa_pr function
source("functions/pa_pr_fun.R")
# source beta_tbi_func function
source("functions/beta_tbi_func.R")
# source est_plot_func function
source("functions/est_plot_func.R")

# adapted occAssess::assessSpatialCov to allow for base layers
source("functions/createData.R") # needed for adapted assessSpatialCov function
source("functions/assessSpatialCov_adapt.R")

# load_rdata function
# loads an RData file, and assigns it to an object name
load_rdata <- function(fileName) {
  load(fileName)
  get(ls()[ls() != "fileName"])
}

# global parameters
startyear <- 1990
endyear <- 2018
ci <- 0.95 # credible interval

```

## Preprocessing ##

Here we preprocess the protected areas data to create a raster of protection for GB

```{r echo = FALSE}

# # grid references
# # load: gr_ref
# gr_ref <- read.csv("data/gr_ref.csv") %>%
#   dplyr::mutate(easting = easting + 500,
#                 northing = northing + 500)

# # british national grid crs
# gbcrs <- "+proj=tmerc +lat_0=49 +lon_0=-2 +k=0.9996012717 +x_0=400000 +y_0=-100000 +ellps=airy +datum=OSGB36 +units=m +no_defs"
# 
# # protected areas (downloaded 13/10/2020 - https://www.protectedplanet.net/country/GBR)
# 
# # load protected area shapefiles
# pa0 <- sf::st_read("data/wdpa_0/WDPA_WDOECM_GBR_shp-polygons.shp") # too big to upload
# pa1 <- sf::st_read("data/wdpa_1/WDPA_WDOECM_GBR_shp-polygons.shp") # too big to upload
# pa2 <- sf::st_read("data/wdpa_2/WDPA_WDOECM_GBR_shp-polygons.shp") # too big to upload
# 
# pa <- rbind(pa0, pa1, pa2)
# 
# # load: pa
# pa <- pa %>%
#   # terrestrial only
#   dplyr::filter(MARINE == 0) %>%
#   # tier 1 sites only
#   dplyr::filter(DESIG_ENG %in% c("National Nature Reserve",
#                              "Ramsar Site, Wetland of International Importance",
#                              "Local Nature Reserve",
#                              "Site Of Special Scientific Interest (Gb)",
#                              "Area Of Special Scientific Interest (Ni)",
#                              "Special Protection Area (Birds Directive)",
#                              "Site of Community Importance (Habitats Directive)",
#                              "Nature Reserve")) %>%
#   # british national grid
#   sf::st_transform(crs = gbcrs)
# 
# # template raster - 1km
# # grid reference spatial locations
# rast_temp <- raster::rasterFromXYZ(dplyr::select(gr_ref, x = easting, y = northing), crs = crs(pa))
# 
# # pa coverage at start of time period
# pa_pr_1990 <- pa_pr(pa = pa, rast_temp = rast_temp, gr_ref = gr_ref, yr = 1990)
# 
# 
# # pa coverage at end of time period
# pa_pr_2018 <- pa_pr(pa = pa, rast_temp = rast_temp, gr_ref = gr_ref, yr = 2018)

```

## Data ##

Here we load the protected area data and classify sites as protected or unprotected

```{r}

# # protected area status
# 
# pa_pr_1990 <- pa_pr_1990 %>%
#   # filter out NAs
#   dplyr::filter(!is.na(grid_ref))
#
# pa_pr_2018 <- pa_pr_2018 %>%
#   # filter out NAs
#   dplyr::filter(!is.na(grid_ref))
# 
# # protected sites
# pro <- pa_pr_1990 %>%
#   # 10 % protection threshold
#   dplyr::filter(pa_prop > 0.1) %>%
#   # protected or not
#   dplyr::mutate(prot = "pa")
# 
# # unprotected sites
# unp <- pa_pr_2018 %>%
#   # 1 % protection threshold
#   dplyr::filter(pa_prop < 0.01) %>%
#   # protected or not
#   dplyr::mutate(prot = "unp")
# 
# # protected status
# pro_stat <- dplyr::bind_rows(pro, unp)
# 
# # save: pro_stat
# saveRDS(pro_stat, "data/pro_stat.rds")

```

# Terrestrial occupancy

### Ants, Bees, E&D, Hoverflies, Ladybirds, Spiders, Wasps

First we extract the occupancy data for all taxonomic groups we are interested in

```{r}

# Taxonomic group(s)
taxa <- c("Ants", "Bees", "Hoverflies", "Ladybirds", "Spiders", "Wasps")
ntax <- length(taxa)

# regions <- c("GB", "pa", "unp")
# nregions <- length(regions)
# 
# vers <- c("2021_Francesca_bwars_rerun", "2021_Francesca_bwars_rerun", "2021_Francesca", "2021_Francesca", "2021_Francesca_marlog_rerun", "2021_Francesca_bwars_rerun")
# 
# # create roster - run for multiple groups & multiple regions
# # ignore warning
# roster <- wrappeR::createRoster(index = 1:(nregions * ntax),
#                        modPath = "/data-s3/occmods/",
#                        metaPath = "/data-s3/metadata/",
#                        ver = rep(vers, each = nregions),
#                        indicator = "all",
#                        region = rep(regions, times = ntax),
#                        nSamps = 999,
#                        minObs = 1,
#                        scaleObs = "region",
#                        write = TRUE,
#                        outPath = "~/proarea/data/filtered/",
#                        group = rep(taxa, each = nregions),
#                        t0 = startyear,
#                        tn = endyear)
# 
# # filtered data
# filt_tax <- lapply(roster, wrappeR::applySamp, parallel = FALSE)

```

## Model threshold quality

```{r}

# ## observation metadata
# 
# meta_pa <- lapply(taxa, function(x) {
#   
#   load_rdata(paste0("data/filtered/", x, "_all_pa_samp.rdata")) %>% 
#     .[[2]] %>% 
#     dplyr::mutate(tax_grp = x)
#   
# }) %>% 
#   dplyr::bind_rows(.) 
# 
# pass_keep <- meta_pa %>% 
#   dplyr::filter(rot_EqualWt_r_pa == TRUE) %>% 
#   .$Species_r_pa

```

## Occupancy estimates

Here we read in prepared occupancy data

```{r}

# # read in filtered data
# occ_read <- function(x, reg) {
#   
#   load_rdata(paste0("data/filtered/", x, "_all_", reg, "_samp.rdata")) %>% 
#     .[[1]] %>% 
#     dplyr::mutate(grp = x)
#   
# }
# 
# # focal years 1990 to 2018
# foc_yrs <- paste0("year_", startyear:endyear)
# 
# ## pa
# 
# occ_pa_prep <- lapply(taxa, function(x) occ_read(x = x, reg = "pa")) %>%
#   dplyr::bind_rows(.) %>%
#   # trim to focal years 1990 to 2018
#   dplyr::select(foc_yrs, iteration, species, grp) %>%
#   # add region aggregate name
#   dplyr::mutate(prot = "pa")
# 
# spp_pa <- data.frame(spp = unique(occ_pa_prep$species))
# 
# ## unp
# 
# occ_unp_prep <- lapply(taxa, function(x) occ_read(x = x, reg = "unp")) %>%
#   dplyr::bind_rows(.) %>%
#   # trim to focal years 1990 to 2018
#   dplyr::select(foc_yrs, iteration, species, grp)  %>%
#   # add region aggregate name
#   dplyr::mutate(prot = "unp")
# 
# spp_unp <- data.frame(spp = unique(occ_unp_prep$species))
# 
# # unify protected and unprotected species lists
# 
# all_spp <- dplyr::inner_join(spp_pa, spp_unp) %>% 
#   # remove non-predatory ladybirds
#   dplyr::filter(!spp %in% c("halyzia sedecimguttata", "henosepilachna argus", "psyllobora vigintiduopunctata", "subcoccinella vigintiquattuorpunctata", "tytthaspis sedecimpunctata")) %>% 
#   # remove non-natives
#   dplyr::filter(!spp %in% c("harmonia axyridis", "steatoda nobilis"))
# 
# # update to match species lists
# occ_pa_prep <- dplyr::filter(occ_pa_prep, species %in% all_spp$spp)
# occ_unp_prep <- dplyr::filter(occ_unp_prep, species %in% all_spp$spp)
# 
# # remove species that didn't pass model quality threshold
# occ_pa_prep <- dplyr::filter(occ_pa_prep, species %in% pass_keep)
# occ_unp_prep <- dplyr::filter(occ_unp_prep, species %in% pass_keep)
# 
# GB_spp <- occ_pa_prep %>% 
#   dplyr::distinct(species, .keep_all = TRUE) %>% 
#   dplyr::count(grp) %>% 
#   tibble::add_row(grp = "Overall", n = sum(.$n))
# 
# occ_GB <- lapply(taxa, function(x) occ_read(x = x, reg = "GB")) %>%
#   dplyr::bind_rows(.) %>%
#   # trim to focal years 1990 to 2018
#   dplyr::select(foc_yrs, iteration, species, grp)  %>%
#   # add region aggregate name
#   dplyr::mutate(prot = "GB") %>% 
#   # match species list
#   dplyr::filter(species %in% occ_pa_prep$species) 

```

# Functional groups

```{r}

occ_pa_prep <- readRDS("data/occ_pa_prep.rds")
occ_unp_prep <- readRDS("data/occ_unp_prep.rds")

# duplicate hoverflies as there are primary contributors to two ecosystem functions
hov_pa <- occ_pa_prep %>% 
  dplyr::filter(grp == "Hoverflies") %>% 
  dplyr::mutate(grp = "Hoverflies2")

hov_unp <- occ_unp_prep %>% 
  dplyr::filter(grp == "Hoverflies") %>% 
  dplyr::mutate(grp = "Hoverflies2")

# ecosystem functions
func_lookup <- data.frame(grp = c("Ants", "Hoverflies", "Ladybirds", "Spiders", "Wasps", "Bees", "Hoverflies2"), func = c(rep("Predators", 5), rep("Pollinators", 2)))

occ_pa_func <- occ_pa_prep %>% 
  dplyr::bind_rows(hov_pa) %>% 
  dplyr::left_join(func_lookup, by = "grp") %>% 
  dplyr::rename(tax_grp = grp, grp = func)

occ_unp_func <- occ_unp_prep %>% 
  dplyr::bind_rows(hov_unp) %>% 
  dplyr::left_join(func_lookup, by = "grp") %>% 
  dplyr::rename(tax_grp = grp, grp = func)

occ_pa <- dplyr::bind_rows(occ_pa_func, occ_pa_prep %>% dplyr::mutate(grp = "Overall"))

occ_unp <- dplyr::bind_rows(occ_unp_func, occ_unp_prep %>% dplyr::mutate(grp = "Overall"))

GB_func_spp <- occ_pa %>% 
  dplyr::distinct(species, grp, .keep_all = TRUE) %>% 
  dplyr::count(grp)

# tidy up
rm(hov_pa, hov_unp, occ_pa_func, occ_pa_prep, occ_unp_func, occ_unp_prep); gc()

```

# Species richness

Here we calculate species richness across and per year

```{r}

# function to calculate species richness per year and across years
sprich <- function(occ_df) {
  
  spr_df <- occ_df %>%
    tidyr::gather(year, occ, dplyr::starts_with("year_")) %>%
    dplyr::group_by(year, grp, iteration, prot) %>%
    # species richness per year
    dplyr::summarise(spr_yr = sum(occ, na.rm = TRUE)) %>%
    dplyr::group_by(grp, iteration, prot) %>%
    # mean species richness across years
    dplyr::summarise(spr = mean(spr_yr, na.rm = TRUE)) %>% 
    # tidy protected area status
    dplyr::mutate(prot = dplyr::recode(prot, pa = "Protected", unp = "Unprotected")) %>% 
    dplyr::mutate(prot = factor(prot, levels = c("Unprotected", "Protected")))
  
}

spr_pa <- sprich(occ_df = occ_pa) %>% 
  dplyr::left_join(GB_func_spp, by = "grp") %>% 
  dplyr::mutate(pspr = spr / n)
spr_unp <- sprich(occ_df = occ_unp) %>% 
  dplyr::left_join(GB_func_spp, by = "grp") %>% 
  dplyr::mutate(pspr = spr / n)

spr_comb <- dplyr::bind_rows(dplyr::select(spr_pa, grp, iteration, prot, spr, pspr), dplyr::select(spr_unp, grp, iteration, prot, spr, pspr))

sprich_diff <- function(spr_df_pa, spr_df_unp) {
  
  spr_df <- dplyr::left_join(dplyr::select(spr_df_pa, grp, iteration, spr, pspr), dplyr::select(spr_df_unp, grp, iteration, spr, pspr), by = c("grp", "iteration")) %>% 
    dplyr::mutate(spr_diff = spr.x - spr.y,
                  pspr_diff = pspr.x - pspr.y)
  
  # average across iterations
  spr_av <- spr_df %>% 
    dplyr::group_by(grp) %>% 
    dplyr::summarise_at(vars(spr_diff, pspr_diff), list(
      ~mean, 
      ~median, 
      low_ci = ~HDInterval::hdi(., credMass = ci)[[1]],
      upp_ci = ~HDInterval::hdi(., credMass = ci)[[2]])) %>% 
    # add significance identifier
    dplyr::mutate(spr_sig = dplyr::case_when(
      spr_diff_low_ci > 0 & spr_diff_upp_ci > 0 ~ "spos",
      spr_diff_low_ci < 0 & spr_diff_upp_ci < 0 ~ "sneg",
      TRUE ~ "ns"
    )) %>% 
    dplyr::mutate(pspr_sig = dplyr::case_when(
      pspr_diff_low_ci > 0 & pspr_diff_upp_ci > 0 ~ "spos",
      pspr_diff_low_ci < 0 & pspr_diff_upp_ci < 0 ~ "sneg",
      TRUE ~ "ns"
    )) %>% 
    dplyr::mutate(spr_sig = factor(spr_sig, levels = c("sneg", "ns", "spos"))) %>% 
    dplyr::mutate(pspr_sig = factor(pspr_sig, levels = c("sneg", "ns", "spos")))
  
  spr_df <- dplyr::left_join(spr_df, dplyr::select(spr_av, grp, spr_sig, pspr_sig), by = c("grp"))
  
  return(list(spr_df, spr_av))
  
}

spr_diff_out <- sprich_diff(spr_df_pa = spr_pa, spr_df_unp = spr_unp)

spr_over <- est_plot_func(grp = "Overall", est_df = spr_comb, est_diff = spr_diff_out, vari = "spr", lim = c(-36, 36), axlab = "Species richness")

# View(spr_over[[2]])
# View(spr_over[[3]])
# View(spr_over[[4]])

# (spr_over[[2]][2,2] - spr_over[[2]][1,2])/spr_over[[2]][1,2] * 100

spr_poll <- est_plot_func(grp = "Pollinators", est_df = spr_comb, est_diff = spr_diff_out, vari = "spr", lim = c(-14, 14), axlab = "Species richness")

spr_pred <- est_plot_func(grp = "Predators", est_df = spr_comb, est_diff = spr_diff_out, vari = "spr", lim = c(-31, 31), axlab = "Species richness")

# # save plots
# cowplot::save_plot("outputs/sprich_over.png", spr_over[[1]], base_height = 6, base_width = 8, dpi = 300)

```

# Multi-species trends

Here we caculate geometric mean occupancy

```{r}

# log of 0 is undefined
nudgeOcc <- function(x, nudgeFac = 0.0001) {
  x[x == 0] <- nudgeFac
  return(x)
}

# function to calculate geometric mean
geomean <- function(x) exp(mean(log(nudgeOcc(x))))

# function to calculate multi-species indicator per iteration
msi <- function(occ_df) {
  
  msi_df <- occ_df %>%
    tidyr::gather(year, occ, dplyr::starts_with("year_")) %>%
    dplyr::group_by(year, grp, iteration, prot) %>%
    # geometric mean
    dplyr::summarise(gm = geomean(occ)) %>% 
    # tidy protected area status
    dplyr::mutate(prot = dplyr::recode(prot, pa = "Protected", unp = "Unprotected")) %>% 
    dplyr::mutate(prot = factor(prot, levels = c("Protected", "Unprotected"))) %>% 
    dplyr::ungroup() %>% 
    dplyr::mutate(year = as.numeric(gsub("year_", "", year)))
  
}

msi_pa <- msi(occ_df = occ_pa)
msi_unp <- msi(occ_df = occ_unp)

msi_sum <- function(msi_df) {
  
  # average across iterations
  msi_av <- msi_df %>% 
    dplyr::group_by(prot, grp, year) %>% 
    dplyr::summarise_at(vars(gm), list(
      ~mean, 
      ~median, 
      #med_low_ci = ~HDInterval::hdi(., credMass = ci_w)[[1]], 
      #med_upp_ci = ~HDInterval::hdi(., credMass = ci_w)[[2]],
      low_ci = ~HDInterval::hdi(., credMass = ci)[[1]],
      upp_ci = ~HDInterval::hdi(., credMass = ci)[[2]]))
}

msi_sum_pa <- msi_sum(msi_df = msi_pa)
msi_sum_unp <- msi_sum(msi_df = msi_unp)

msi_sum_comb <- dplyr::bind_rows(msi_sum_pa, msi_sum_unp)

trend_plot_func <- function(grp, msi_sum, lim = NULL, leg = FALSE, leg_coord = NULL, axlab = "Geometric mean occupancy") {
  
  low_lim <- lim[[1]]
  upp_lim <- lim[[2]]
  
  numb <- GB_func_spp %>% 
    dplyr::filter(grp == !!grp)
  
  msi <- msi_sum %>% 
    dplyr::filter(grp == !!grp)
  
  refe <- dplyr::filter(msi, year == 1990)
  
  trend_plot <- ggplot(data = msi, aes(x = year, y = median, group = prot)) +
    geom_hline(data = refe, aes(colour = prot, yintercept = median), lty = 2) +
    # 95% CrI
    geom_ribbon(aes(ymin = low_ci, ymax = upp_ci, fill = prot), alpha = 0.6) +
    geom_line(aes(y = median, colour = prot)) +
    geom_point(aes(colour = prot), size = 0.6) +
    # scales
    scale_y_continuous(name = axlab, limits = c(low_lim, upp_lim)) +
    scale_x_continuous(name = "Year", breaks = scales::pretty_breaks(6), expand = c(0,0)) +
    scale_colour_manual(name = "", values = c("#0077BB", "#EE7733")) +
    scale_fill_manual(name = "", values = c("#5DC4FF", "#F6BB99")) +
    theme(legend.position = "none")
  
  if(leg == TRUE) trend_plot <- trend_plot + 
    annotate("text", x = leg_coord[1], y = leg_coord[2], label = "Protected", colour = "#0077BB", size = 6) + 
    annotate("text", x = leg_coord[3], y = leg_coord[4], label = "Unprotected", colour = "#EE7733", size = 6) 
  
  # trend_title <- cowplot::ggdraw() +
  #   cowplot::draw_label(paste0(grp, " (n = ", format(numb$n, big.mark = ","), ")"), fontface = 'bold')
  # 
  # trend_plot <- cowplot::plot_grid(trend_title, trend_plot, nrow = 2, rel_heights = c(0.08, 1))
  
  return(trend_plot)
  
}

msi_over <- trend_plot_func(grp = "Overall", msi_sum = msi_sum_comb, lim = c(0.043, 0.082), leg = TRUE, leg_coord = c(2015, 0.076, 2015, 0.053))

msi_poll <- trend_plot_func(grp = "Pollinators", msi_sum = msi_sum_comb, lim = c(0.037, 0.087), axlab = "Geometric\nmean occupancy")

msi_pred <- trend_plot_func(grp = "Predators", msi_sum = msi_sum_comb, lim = c(0.044, 0.084), axlab = "Geometric\nmean occupancy")

```

# Trend change

Here we calculate annual growth rate between the first and last year

```{r}

# function to calculate annual growth rate per species
trend_change <- function(occ_df, taxa) {
  
  out <- lapply(taxa, function(x) {
    
    spp_df <- occ_df %>% 
      dplyr::filter(grp == x) %>% 
      # only non-na columns
      dplyr::select(which(colSums(is.na(.)) < 1)) %>% 
      # select first and last year
      dplyr::select(1, (ncol(.) - 4), iteration, species, prot, grp) %>%
      # record first year
      dplyr::mutate(first_year = colnames(.)[1]) %>%
      dplyr::mutate(first_year = as.numeric(gsub("year_", "", first_year))) %>%
      # record last year
      dplyr::mutate(last_year = colnames(.)[2]) %>%
      dplyr::mutate(last_year = as.numeric(gsub("year_", "", last_year))) %>%
      # set column names
      dplyr::rename(occ_first = 1, occ_last = 2) %>% 
      # number of years
      dplyr::mutate(nyr = last_year - first_year) %>%
      # replace zero with very small number - can't divide by zero
      dplyr::mutate(occ_first = ifelse(occ_first == 0, 1e-07, occ_first)) %>%
      # annual growth rate - see occurrenceChange()
      dplyr::mutate(change = (((occ_last / occ_first) ^ (1 / nyr)) - 1) * 100)
    
  }) %>% 
    # bind across taxonomic groups
    dplyr::bind_rows(.)
  
}

trend_pa <- trend_change(occ_df = dplyr::select(occ_pa, -tax_grp), taxa = unique(occ_unp$grp))
trend_unp <- trend_change(occ_df = dplyr::select(occ_unp, -tax_grp), taxa = unique(occ_unp$grp))

# function to calculate annual growth rate per group
tax_change <- function(trend_df) {
  
  chng_tax <- trend_df %>% 
    dplyr::group_by(grp, iteration) %>% 
    dplyr::summarise(change = mean(change))
}

chng_pa <- tax_change(trend_df = trend_pa)
chng_unp <- tax_change(trend_df = trend_unp)

chng_comb <- dplyr::bind_rows(
  dplyr::select(chng_pa, grp, iteration, change) %>% 
    dplyr::mutate(prot = "Protected"), 
  dplyr::select(chng_unp, grp, iteration, change) %>% 
    dplyr::mutate(prot = "Unprotected")) %>% 
    dplyr::mutate(prot = factor(prot, levels = c("Unprotected", "Protected")))

change_diff <- function(chng_df_pa, chng_df_unp) {
  
  # combine protected and unprotected
  chng_df <- dplyr::left_join(chng_df_pa, chng_df_unp, by = c("grp", "iteration")) %>% 
    dplyr::rename(change_pa = change.x, change_unp = change.y) %>% 
    dplyr::mutate(change_diff = change_pa - change_unp) %>% 
    dplyr::ungroup()
  
  # average across iterations
  chng_av <- chng_df %>% 
    dplyr::group_by(grp) %>% 
    dplyr::summarise_at(vars(change_diff), list(
      change_diff_mean = ~mean, 
      change_diff_median = ~median, 
      # change_diff_med_low_ci = ~HDInterval::hdi(., credMass = ci_w)[[1]], 
      # change_diff_med_upp_ci = ~HDInterval::hdi(., credMass = ci_w)[[2]],
      change_diff_low_ci = ~HDInterval::hdi(., credMass = ci)[[1]],
      change_diff_upp_ci = ~HDInterval::hdi(., credMass = ci)[[2]])) %>% 
    # add significance identifier
    dplyr::mutate(change_sig = dplyr::case_when(
      change_diff_low_ci > 0 & change_diff_upp_ci > 0 ~ "spos",
      change_diff_low_ci < 0 & change_diff_upp_ci < 0 ~ "sneg",
      # change_diff_med_low_ci > 0 & change_diff_med_upp_ci > 0 ~ "wpos",
      # change_diff_med_low_ci > 0 & change_diff_med_upp_ci > 0 ~ "wneg",
      TRUE ~ "ns"
    )) %>% 
    dplyr::mutate(change_sig = factor(change_sig, levels = c("sneg", "ns", "spos")))
  
  chng_df <- dplyr::left_join(chng_df, dplyr::select(chng_av, grp, change_sig), by = c("grp"))
  
    return(list(chng_df, chng_av))
}

chng_diff_out <- change_diff(chng_df_pa = chng_pa, chng_df_unp = chng_unp)

chng_over <- est_plot_func(grp = "Overall", est_df = chng_comb, est_diff = chng_diff_out, vari = "change", lim = c(-1.8, 1.8), axlab = "Growth rate")

# View(chng_over[[2]])
# View(chng_over[[3]])
# View(chng_over[[4]])

chng_poll <- est_plot_func(grp = "Pollinators", est_df = chng_comb, est_diff = chng_diff_out, vari = "change", lim = c(-2.8, 2.8), axlab = "Growth rate")

chng_pred <- est_plot_func(grp = "Predators", est_df = chng_comb, est_diff = chng_diff_out, vari = "change", lim = c(-2.4, 2.4), axlab = "Growth rate")

chng_comb_over <- cowplot::plot_grid(msi_over, chng_over[[1]], nrow = 2, labels = "AUTO")

# # save plot
# cowplot::save_plot("outputs/chng_comb_over.png", chng_comb_over, base_height = 11, base_width = 9, dpi = 300)

```

# Temporal beta diversity

```{r}

# beta_tbi <- lapply(unique(occ_pa$grp), function(xgrp) {
# 
#   out <- lapply(1:999, function(x) beta_tbi_func(grp = xgrp, it = x, occ_df_pa = occ_pa, occ_df_unp = occ_unp) %>%
#     # bind across iterations
#   dplyr::bind_rows(.))
# 
# }) %>%
#   # bind across groups
#   dplyr::bind_rows()
# 
# saveRDS(beta_tbi, "data/beta_tbi.rds")

beta_tbi <- readRDS("data/beta_tbi.rds") %>% 
  dplyr::mutate(prot = factor(prot, levels = c("Unprotected", "Protected")))

beta_contr <- dplyr::left_join(dplyr::filter(beta_tbi, prot == "Protected"), dplyr::filter(beta_tbi, prot == "Unprotected"), by = c("iteration", "grp"))

beta_diff <- function(beta_df) {
  
  beta_df <- beta_df %>% 
    dplyr::mutate(tbi_val_diff = tbi_val.x - tbi_val.y,
                  loss_diff = loss.x - loss.y,
                  gain_diff = gain.x - gain.y)
  
  # average across iterations
  beta_av <- beta_df %>% 
    dplyr::group_by(grp) %>% 
    dplyr::summarise_at(vars(dplyr::ends_with("diff")), list(
      ~mean, 
      ~median, 
      low_ci = ~HDInterval::hdi(., credMass = ci)[[1]],
      upp_ci = ~HDInterval::hdi(., credMass = ci)[[2]])) %>% 
    # add significance identifier
    dplyr::mutate(tbi_val_sig = dplyr::case_when(
      tbi_val_diff_low_ci > 0 & tbi_val_diff_upp_ci > 0 ~ "spos",
      tbi_val_diff_low_ci < 0 & tbi_val_diff_upp_ci < 0 ~ "sneg",
      TRUE ~ "ns"
    )) %>% 
    dplyr::mutate(loss_sig = dplyr::case_when(
      loss_diff_low_ci > 0 & loss_diff_upp_ci > 0 ~ "spos",
      loss_diff_low_ci < 0 & loss_diff_upp_ci < 0 ~ "sneg",
      TRUE ~ "ns"
    )) %>% 
    dplyr::mutate(gain_sig = dplyr::case_when(
      gain_diff_low_ci > 0 & gain_diff_upp_ci > 0 ~ "spos",
      gain_diff_low_ci < 0 & gain_diff_upp_ci < 0 ~ "sneg",
      TRUE ~ "ns"
    )) %>% 
    dplyr::mutate(tbi_val_sig = factor(tbi_val_sig, levels = c("sneg", "ns", "spos")),
                  loss_sig = factor(loss_sig, levels = c("sneg", "ns", "spos")),
                  gain_sig = factor(gain_sig, levels = c("sneg", "ns", "spos")))
  
  beta_df <- dplyr::left_join(beta_df, dplyr::select(beta_av, grp, tbi_val_sig, loss_sig, gain_sig), by = c("grp"))
  
    return(list(beta_df, beta_av))
}

beta_diff_out <- beta_diff(beta_df = beta_contr)

# total

beta_over <- est_plot_func(grp = "Overall", est_df = beta_tbi, est_diff = beta_diff_out, vari = "tbi_val", lim = c(-0.022, 0.022), axlab = "Temporal beta diversity")

# View(beta_over[[2]])
# View(beta_over[[3]])
# View(beta_over[[4]])

beta_poll <- est_plot_func(grp = "Pollinators", est_df = beta_tbi, est_diff = beta_diff_out, vari = "tbi_val", lim = c(-0.031, 0.031), axlab = "Temporal\nbeta diversity")

beta_pred <- est_plot_func(grp = "Predators", est_df = beta_tbi, est_diff = beta_diff_out, vari = "tbi_val", lim = c(-0.025, 0.025), axlab = "Temporal\nbeta diversity")

# loss

loss_over <- est_plot_func(grp = "Overall", est_df = beta_tbi, est_diff = beta_diff_out, vari = "loss", lim = c(-0.027, 0.027), axlab = "Loss")

loss_poll <- est_plot_func(grp = "Pollinators", est_df = beta_tbi, est_diff = beta_diff_out, vari = "loss", lim = c(-0.023, 0.023), axlab = "Loss")

loss_pred <- est_plot_func(grp = "Predators", est_df = beta_tbi, est_diff = beta_diff_out, vari = "loss", lim = c(-0.038, 0.038), axlab = "Loss")

# gain

gain_over <- est_plot_func(grp = "Overall", est_df = beta_tbi, est_diff = beta_diff_out, vari = "gain", lim = c(-0.024, 0.024), axlab = "Gain")

gain_poll <- est_plot_func(grp = "Pollinators", est_df = beta_tbi, est_diff = beta_diff_out, vari = "gain", lim = c(-0.031, 0.031), axlab = "Gain")

gain_pred <- est_plot_func(grp = "Predators", est_df = beta_tbi, est_diff = beta_diff_out, vari = "gain", lim = c(-0.031, 0.031), axlab = "Gain")

beta_over_plot <- cowplot::plot_grid(NULL, beta_over[[1]], NULL, nrow = 1, rel_widths = c(0.4, 1, 0.4), labels = c("", "A", "")) %>% cowplot::plot_grid(., cowplot::plot_grid(loss_over[[1]], gain_over[[1]], nrow = 1, labels = c("B", "C")), nrow = 2)

# # save plot
# cowplot::save_plot("outputs/beta_over.png", beta_over_plot, base_height = 8, base_width = 12, dpi = 300)

#### Pollinators ####

beta_poll_plot <- cowplot::plot_grid(NULL, beta_poll[[1]], NULL, nrow = 1, rel_widths = c(0.4, 1, 0.4), labels = c("", "D", "")) %>% cowplot::plot_grid(., cowplot::plot_grid(loss_poll[[1]], gain_poll[[1]], nrow = 1, labels = c("E", "F")), nrow = 2)

poll_plot <- cowplot::plot_grid(cowplot::plot_grid(NULL, spr_poll[[1]], NULL, labels = c("", "A", ""), rel_widths = c(0.4, 1, 0.4), ncol = 3), cowplot::plot_grid(msi_poll, chng_poll[[1]], labels = c("B", "C"), ncol = 2), beta_poll_plot, nrow = 3, rel_heights = c(1, 1, 2))

# # save plot
# cowplot::save_plot("outputs/poll.png", poll_plot, base_height = 12, base_width = 10, dpi = 300)

beta_pred_plot <- cowplot::plot_grid(NULL, beta_pred[[1]], NULL, nrow = 1, rel_widths = c(0.4, 1, 0.4), labels = c("", "D", "")) %>% cowplot::plot_grid(., cowplot::plot_grid(loss_pred[[1]], gain_pred[[1]], nrow = 1, labels = c("E", "F")), nrow = 2)

pred_plot <- cowplot::plot_grid(cowplot::plot_grid(NULL, spr_pred[[1]], NULL, labels = c("", "A", ""), rel_widths = c(0.4, 1, 0.4), ncol = 3), cowplot::plot_grid(msi_pred, chng_pred[[1]], labels = c("B", "C"), ncol = 2), beta_pred_plot, nrow = 3, rel_heights = c(1, 1, 2))

# # save plot
# cowplot::save_plot("outputs/pred.png", pred_plot, base_height = 12, base_width = 10, dpi = 300)

```

## GB plot

```{r}

# fig 1
# plot of protected vs unprotected grid cells

# sites used
sites_orig <- readRDS("data/regions_non-crossed.rds")

sites <- sites_orig %>% 
  # exclude Northern Ireland
  dplyr::filter(NORTHERN_IRELAND == 0) %>% 
  dplyr::mutate(prot = ifelse(pa == 1 & unp == 0, "pa", 
                              ifelse(pa == 0 & unp == 1, "unp", "other"))) %>% 
  # order levels - pa first
  dplyr::mutate(prot = factor(prot, levels = c("pa", "unp", "other"))) %>%
  dplyr::select(grid_ref, prot)

# # convert grid references to easting and northing
# grcoord <- BRCmap::gr2sp_points(sites$grid_ref) %>%
#   as.data.frame(.)
# 
# sites_gr <- dplyr::left_join(sites, grcoord, by = c("grid_ref" = "GRIDREF"))

sites_gr <- readRDS("data/sites_gr.rds")
  
pro_plot <- ggplot2::ggplot(data = sites_gr, aes(x = EASTING, y = NORTHING, fill = prot)) +
  geom_raster() +
  coord_equal() +
  scale_fill_manual(breaks = c("pa", "unp"), labels = c("Protected", "Unprotected"), values = c("#0077BB", "#EE7733", "grey")) +
  theme_void() +
  theme(legend.title = element_blank(),
        legend.position = "bottom")

euro <- map_data("world") %>% 
  dplyr::mutate(GB = as.factor(ifelse(subregion %in% c("Great Britain", "Wales", "Isle of Wight", "Scotland"), 1, 0)))

euro_plot <- ggplot(euro, aes(x = long, y = lat)) +
  geom_polygon(aes(group = group, fill = GB, colour = GB), size = 0.3) +
  coord_map(project = "orthographic", xlim = c(-12,44), ylim = c(35,70)) +
  scale_fill_manual(values = c("grey", "black")) +
  scale_colour_manual(values = c("darkgrey", "black")) +
  scale_x_continuous(breaks = seq(-140, 60, by = 20))+
  scale_y_continuous(breaks = seq(10, 90, by = 20)) +
  theme_bw() +
  theme(panel.grid.major = element_line(colour = "lightgrey"),
        panel.border = element_rect(colour = "black"),
        axis.title = element_blank(),
        axis.ticks = element_blank(),
        axis.text = element_blank(),
        legend.position = "none")

pro_sub_plot <- cowplot::ggdraw(pro_plot) +
  cowplot::draw_plot(euro_plot, 0.63, 0.55, 0.35, 0.35)

# # save: pro_sub_plot
# cowplot::save_plot("outputs/pro_sub.png", pro_sub_plot, base_height = 6, base_width = 6, dpi = 300)

```

## Number of records

```{r}

# meta_pa <- lapply(taxa, function(x) {
# 
#   load_rdata(paste0("data/filtered/", x, "_all_pa_samp.rdata")) %>%
#     .[[2]] %>%
#     dplyr::mutate(tax_grp = x)
# 
# }) %>%
#   dplyr::bind_rows(.) %>%
#   dplyr::filter(Species_r_pa %in% occ_pa$species)
# 
# rec_pa <- meta_pa %>% 
#   dplyr::group_by(tax_grp) %>% 
#   dplyr::summarise(recs = sum(n_obs_regional_r_pa))
# 
# meta_unp <- lapply(taxa, function(x) {
# 
#   load_rdata(paste0("data/filtered/", x, "_all_unp_samp.rdata")) %>%
#     .[[2]] %>%
#     dplyr::mutate(tax_grp = x)
# 
# }) %>%
#   dplyr::bind_rows(.) %>%
#   dplyr::filter(Species_r_unp %in% occ_unp$species)
# 
# rec_unp <- meta_unp %>%
#   dplyr::group_by(tax_grp) %>%
#   dplyr::summarise(recs = sum(n_obs_regional_r_unp))
# 
# rec_all <- dplyr::bind_cols(rec_pa, rec_unp) %>% 
#   dplyr::mutate(rec_all = recs + recs1)
# 
# saveRDS(rec_all, "data/rec_all.rds")

rec_all <- readRDS("data/rec_all.rds")

sum(rec_all$rec_all)

```

## Preprocessing input data for occAssess

```{r}

# # function to convert first letter to uppercase
# firstup <- function(x) {
#   substr(x, 1, 1) <- toupper(substr(x, 1, 1))
#   x
# }
# 
# # function to tidy observations used in models
# # occ_pa and sites_gr needed
# obs <- function(tax_path, tax_grp, filetype, chained = FALSE, min_iter) {
#   
#   spp <- dplyr::filter(occ_pa, tax_grp == !!tax_grp)$species %>% 
#     unique(.) %>% 
#     firstup(.)
#   
#   all_out <- lapply(spp, function(sp) {
#     
#     if(isTRUE(chained)) {
#       
#       out_meta <- load_rdata(paste0(tax_path, sp, "_", min_iter, "_1.rdata"))
#       
#     } else {
#       
#       if(filetype == "rds") {
#         
#         out_meta <- readRDS(paste0(tax_path, sp, ".rds"))
#         
#         }
#     
#       else if(filetype == "rdata") {
#       
#         out_meta <- load_rdata(paste0(tax_path, sp, ".rdata"))
#       
#       }
#     }
#     
#     dat <- out_meta$model$data()
#     
#     dat_df <- data.frame(year = dat$Year, # year
#                          rec = dat$y, # records
#                          site = dat$Site) # sites
# 
#     # site number to grid ref look-up
#     dat_site <- dat_df %>% 
#       dplyr::distinct(site) %>% 
#       dplyr::mutate(grid = unique(out_meta$sites_included[[1]]))
#     
#     region_site_pa <- dat[[paste0("r_", "pa")]][dat$Site]
#     region_site_unp <- dat[[paste0("r_", "unp")]][dat$Site]
#     
#     # sites included within region
#     dat_reg <- cbind(dat_df, region_site_pa, region_site_unp) %>% 
#       dplyr::left_join(dat_site)
#           
#     # subset to temporal window
#     dat_reg_pa <- dat_reg[dat_reg$region_site_pa == 1 & dat_reg$year >= (startyear - (out_meta$min_year - 1)) & dat_reg$year <= (endyear - (out_meta$min_year - 1)), ] %>% 
#       dplyr::filter(rec == 1) %>% 
#       dplyr::mutate(prot = "pa")
#     
#     dat_reg_unp <- dat_reg[dat_reg$region_site_unp == 1 & dat_reg$year >= (startyear - (out_meta$min_year - 1)) & dat_reg$year <= (endyear - (out_meta$min_year - 1)), ] %>% 
#       dplyr::filter(rec == 1) %>% 
#       dplyr::mutate(prot = "unp")
#     
#     dat_comb <- dplyr::bind_rows(dat_reg_pa, dat_reg_unp) %>% 
#       dplyr::mutate(year = (year + 1970) - 1) %>% 
#       dplyr::left_join(dplyr::select(sites_gr, -prot), by = c("grid" = "grid_ref")) %>% 
#       dplyr::mutate(species = tolower(sp)) %>% 
#       dplyr::mutate(tax_grp = tax_grp) %>% 
#       dplyr::select(species, tax_grp, prot, grid, EASTING, NORTHING, year)
#     
#   }) %>% 
#     dplyr::bind_rows()
#   
#   return(all_out)
#   
# }
# 
# ants_obs <- obs(tax_path = "/data-s3/occmods/Ants/occmod_outputs/2021_Francesca_bwars_rerun/", tax_grp = "Ants", filetype = "rdata")
# 
# # saveRDS(ants_obs, "data/ants_obs.rds")
# 
# bees_obs <- obs(tax_path = "/data-s3/occmods/Bees/occmod_outputs/2021_Francesca_bwars_rerun/", tax_grp = "Bees", filetype = "rdata")
# 
# # saveRDS(bees_obs, "data/bees_obs.rds")
# 
# hover_obs <- obs(tax_path = "/data-s3/occmods/Hoverflies/occmod_outputs/2021_Francesca/", tax_grp = "Hoverflies", filetype = "rdata", chained = TRUE, min_iter = 2000)
# 
# # saveRDS(hover_obs, "data/hover_obs.rds")
# 
# lady_obs <- obs(tax_path = "/data-s3/occmods/Ladybirds/occmod_outputs/2021_Francesca/", tax_grp = "Ladybirds", filetype = "rdata", chained = TRUE, min_iter = 8000)
# 
# # saveRDS(lady_obs, "data/lady_obs.rds")
# 
# spider_obs <- obs(tax_path = "/data-s3/occmods/Spiders/occmod_outputs/2021_Francesca_marlog_rerun/", tax_grp = "Spiders", filetype = "rdata")
# 
# # saveRDS(spider_obs, "data/spider_obs.rds")
# 
# wasps_obs <- obs(tax_path = "/data-s3/occmods/Wasps/occmod_outputs/2021_Francesca_bwars_rerun/", tax_grp = "Wasps", filetype = "rdata")
# 
# # saveRDS(wasps_obs, "data/wasps_obs.rds")
# 
# all_obs <- dplyr::bind_rows(ants_obs, bees_obs, hover_obs, lady_obs, spider_obs, wasps_obs)
# 
# # saveRDS(all_obs, "data/all_obs.rds")

all_obs <- readRDS("data/all_obs.rds")

```

## rhat

```{r}

# rhat <- function(tax_path, tax_grp, filetype, chained = FALSE, max_iter) {
#   
#   spp <- dplyr::filter(occ_pa, tax_grp == !!tax_grp)$species %>% 
#     unique(.) %>% 
#     firstup(.)
#   
#   bugs <- lapply(spp, function(sp) {
#     
#     if(isTRUE(chained)) {
#       
#       out_dat <- load_rdata(paste0(tax_path, sp, "_", max_iter, "_1.rdata"))
#       
#       if(!is.null(out_dat) & is.null(out_dat$model)) out_dat <- out_dat$out
#       
#     } else {
#       
#       if(filetype == "rds") {
#         
#         out_dat <- readRDS(paste0(tax_path, sp, ".rds"))
#         
#         }
#     
#       else if(filetype == "rdata") {
#       
#         out_dat <- load_rdata(paste0(tax_path, sp, ".rdata"))
#       
#       }
#     }
#     
#     bugs <- out_dat$BUGSoutput$summary %>% 
#       data.frame() %>% 
#       tibble::rownames_to_column("para") %>% 
#       dplyr::filter(stringr::str_detect(para, "psi.fs")) %>% 
#       dplyr::filter(!stringr::str_detect(para, "psi.fs.r_ENGLAND|psi.fs.r_GB|psi.fs.r_NORTHERN_IRELAND|psi.fs.r_SCOTLAND|psi.fs.r_UK|psi.fs.r_WALES|psi.fs.r_high|psi.fs.r_low|psi.fs.r_no_agri|psi.fs.r_pa|psi.fs.r_unp")) %>% 
#       dplyr::mutate(species = tolower(sp)) %>% 
#       dplyr::mutate(tax_grp = tax_grp)
#     
#   }) %>% 
#     dplyr::bind_rows()
#   
#   return(bugs)
#   
# }
# 
# ants_rhat <- rhat(tax_path = "/data-s3/occmods/Ants/occmod_outputs/2021_Francesca_bwars_rerun/", tax_grp = "Ants", filetype = "rdata")
# 
# # saveRDS(ants_rhat, "data/ants_rhat.rds")
# 
# bees_rhat <- rhat(tax_path = "/data-s3/occmods/Bees/occmod_outputs/2021_Francesca_bwars_rerun/", tax_grp = "Bees", filetype = "rdata")
# 
# # saveRDS(bees_rhat, "data/bees_rhat.rds")
# 
# hover_rhat <- rhat(tax_path = "/data-s3/occmods/Hoverflies/occmod_outputs/2021_Francesca/", tax_grp = "Hoverflies", filetype = "rdata", chained = TRUE, max_iter = 32000)
# 
# # saveRDS(hover_rhat, "data/hover_rhat.rds")
# 
# lady_rhat <- rhat(tax_path = "/data-s3/occmods/Ladybirds/occmod_outputs/2021_Francesca/", tax_grp = "Ladybirds", filetype = "rdata", chained = TRUE, max_iter = 32000)
# 
# # saveRDS(lady_rhat, "data/lady_rhat.rds")
# 
# spider_rhat <- rhat(tax_path = "/data-s3/occmods/Spiders/occmod_outputs/2021_Francesca_marlog_rerun/", tax_grp = "Spiders", filetype = "rdata")
# 
# # saveRDS(spider_rhat, "data/spider_rhat.rds")
# 
# wasps_rhat <- rhat(tax_path = "/data-s3/occmods/Wasps/occmod_outputs/2021_Francesca_bwars_rerun/", tax_grp = "Wasps", filetype = "rdata")
# 
# # saveRDS(wasps_rhat, "data/wasps_rhat.rds")
# 
# all_rhat <- dplyr::bind_rows(ants_rhat, bees_rhat, hover_rhat, lady_rhat, spider_rhat, wasps_rhat)
# 
# # saveRDS(all_rhat, "data/all_rhat.rds")

all_rhat <- readRDS("data/all_rhat.rds")

# first and last rhats
fl_rhat <- all_rhat %>% 
  dplyr::group_by(species) %>% 
  slice(c(1,n()))

conv_spp <- fl_rhat %>% 
  dplyr::group_by(species) %>%
  dplyr::filter(all(Rhat < 1.1))

```

# Species whose models converged
# Rhat < 1.1 for first and last years

```{r}

occ_pa_conv <- dplyr::filter(occ_pa, species %in% conv_spp$species)
occ_unp_conv <- dplyr::filter(occ_unp, species %in% conv_spp$species)

spr_pa_conv <- sprich(occ_df = occ_pa_conv) %>% 
  dplyr::left_join(GB_func_spp, by = "grp") %>% 
  dplyr::mutate(pspr = spr / n)
spr_unp_conv <- sprich(occ_df = occ_unp_conv) %>% 
  dplyr::left_join(GB_func_spp, by = "grp") %>% 
  dplyr::mutate(pspr = spr / n)

spr_comb_conv <- dplyr::bind_rows(dplyr::select(spr_pa_conv, grp, iteration, prot, spr, pspr), dplyr::select(spr_unp_conv, grp, iteration, prot, spr, pspr))

spr_diff_out_conv <- sprich_diff(spr_df_pa = spr_pa_conv, spr_df_unp = spr_unp_conv)

spr_over_conv <- est_plot_func(grp = "Overall", est_df = spr_comb_conv, est_diff = spr_diff_out_conv, vari = "spr", lim = c(-14, 14), axlab = "Species richness")

msi_pa_conv <- msi(occ_df = occ_pa_conv)
msi_unp_conv <- msi(occ_df = occ_unp_conv)

msi_sum_pa_conv <- msi_sum(msi_df = msi_pa_conv)
msi_sum_unp_conv <- msi_sum(msi_df = msi_unp_conv)

msi_sum_comb_conv <- dplyr::bind_rows(msi_sum_pa_conv, msi_sum_unp_conv)

msi_over_conv <- trend_plot_func(grp = "Overall", msi_sum = msi_sum_comb_conv, lim = c(0.055, 0.12), leg = FALSE, axlab = "Geometric\nmean occupancy")

trend_pa_conv <- trend_change(occ_df = dplyr::select(occ_pa_conv, -tax_grp), taxa = unique(occ_unp_conv$grp))
trend_unp_conv <- trend_change(occ_df = dplyr::select(occ_unp_conv, -tax_grp), taxa = unique(occ_unp_conv$grp))

chng_pa_conv <- tax_change(trend_df = trend_pa_conv)
chng_unp_conv <- tax_change(trend_df = trend_unp_conv)

chng_comb_conv <- dplyr::bind_rows(
  dplyr::select(chng_pa_conv, grp, iteration, change) %>% 
    dplyr::mutate(prot = "Protected"), 
  dplyr::select(chng_unp_conv, grp, iteration, change) %>% 
    dplyr::mutate(prot = "Unprotected")) %>% 
    dplyr::mutate(prot = factor(prot, levels = c("Unprotected", "Protected")))

chng_diff_out_conv <- change_diff(chng_df_pa = chng_pa_conv, chng_df_unp = chng_unp_conv)

chng_over_conv <- est_plot_func(grp = "Overall", est_df = chng_comb_conv, est_diff = chng_diff_out_conv, vari = "change", lim = c(-2.8, 2.8), axlab = "Growth rate")

chng_comb_over_conv <- cowplot::plot_grid(msi_over_conv, chng_over_conv[[1]], nrow = 2, labels = "AUTO")

# beta_tbi_conv <- lapply(unique(occ_pa_conv$grp), function(xgrp) {
# 
#   out <- lapply(1:999, function(x) beta_tbi_func(grp = xgrp, it = x, occ_df_pa = occ_pa_conv, occ_df_unp = occ_unp_conv)) %>%
#     # bind across iterations
#   dplyr::bind_rows()
# 
# }) %>%
#   # bind across groups
#   dplyr::bind_rows()
# 
# saveRDS(beta_tbi_conv, "data/beta_tbi_conv.rds")

beta_tbi_conv <- readRDS("data/beta_tbi_conv.rds") %>% 
  dplyr::mutate(prot = factor(prot, levels = c("Unprotected", "Protected")))

beta_contr_conv <- dplyr::left_join(dplyr::filter(beta_tbi_conv, prot == "Protected"), dplyr::filter(beta_tbi_conv, prot == "Unprotected"), by = c("iteration", "grp"))

beta_diff_out_conv <- beta_diff(beta_df = beta_contr_conv)

# total

beta_over_conv <- est_plot_func(grp = "Overall", est_df = beta_tbi_conv, est_diff = beta_diff_out_conv, vari = "tbi_val", lim = c(-0.028, 0.028), axlab = "Temporal beta diversity")

loss_over_conv <- est_plot_func(grp = "Overall", est_df = beta_tbi_conv, est_diff = beta_diff_out_conv, vari = "loss", lim = c(-0.023, 0.023), axlab = "Loss")

# View(loss_over_conv[[2]])
# View(loss_over_conv[[3]])
# View(loss_over_conv[[4]])

# gain

gain_over_conv <- est_plot_func(grp = "Overall", est_df = beta_tbi_conv, est_diff = beta_diff_out_conv, vari = "gain", lim = c(-0.014, 0.014), axlab = "Gain")

# View(gain_over_conv[[2]])
# View(gain_over_conv[[3]])
# View(gain_over_conv[[4]])

beta_over_plot_conv <- cowplot::plot_grid(NULL, beta_over_conv[[1]], NULL, nrow = 1, rel_widths = c(0.4, 1, 0.4), labels = c("", "A", "")) %>% cowplot::plot_grid(., cowplot::plot_grid(loss_over_conv[[1]], gain_over_conv[[1]], nrow = 1, labels = c("B", "C")), nrow = 2)

beta_over_plot_conv <- cowplot::plot_grid(NULL, beta_over_conv[[1]], NULL, nrow = 1, rel_widths = c(0.4, 1, 0.4), labels = c("", "D", "")) %>% cowplot::plot_grid(., cowplot::plot_grid(loss_over_conv[[1]], gain_over_conv[[1]], nrow = 1, labels = c("E", "F")), nrow = 2)

conv_plot <- cowplot::plot_grid(cowplot::plot_grid(NULL, spr_over_conv[[1]], NULL, labels = c("", "A", ""), rel_widths = c(0.4, 1, 0.4), ncol = 3), cowplot::plot_grid(msi_over_conv, chng_over_conv[[1]], labels = c("B", "C"), ncol = 2), beta_over_plot_conv, nrow = 3, rel_heights = c(1, 1, 2))

# # save plot
# cowplot::save_plot("outputs/conv.png", conv_plot, base_height = 12, base_width = 10, dpi = 300)

```

# Spatiotemporal coverage

```{r}

all_obs <- readRDS("data/occ_assess/all_obs.rds") %>% 
  dplyr::mutate(spatialUncertainty = NA,
                identifier = "Overall")

sites_gr <- readRDS("occ_assess/sites_gr.rds")

pa_rast <- raster::rasterFromXYZ(dplyr::filter(sites_gr, prot == "pa") %>% dplyr::mutate(z = 1) %>% dplyr::select(x = EASTING, y = NORTHING, z))

unp_rast <- raster::rasterFromXYZ(dplyr::filter(sites_gr, prot == "unp") %>% dplyr::mutate(z = 1) %>% dplyr::select(x = EASTING, y = NORTHING, z))

periods <- as.list(1990:2018)
# periods_4 <- list(1990:1994, 1995:1998, 1999:2002, 2003:2006, 2007:2010, 2011:2014, 2015:2018)
periods_7 <- list(1990:1997, 1998:2004, 2005:2011, 2012:2018)

period_names <- as.character(periods_7)

# function to relabel facets
period_labeller <- function(variable,value){
  return(period_names[value])
}

colrs <- viridis::viridis(20)

colrs_nam <- c("1" = colrs[1], "2" = colrs[2], "3" = colrs[3], "4" = colrs[4], "5" = colrs[5], "6" = colrs[6], "7" = colrs[7], "8" = colrs[8], "9" = colrs[9], "10" = colrs[10], "11" = colrs[11], "12" = colrs[12], "13" = colrs[13], "14" = colrs[14], "15" = colrs[15], "16" = colrs[16], "17" = colrs[17], "18" = colrs[18], "19" = colrs[19], "20" = colrs[20], "21" = colrs[21], "22" = colrs[22], "23" = colrs[23], "24" = colrs[24], "25" = colrs[25], "26" = colrs[26], "27" = colrs[27], "28" = colrs[28])

# taxonomic groups

## Spatiotemporal coverage ##

# protected
maps_pers_tax_pa <- assessSpatialCov_adapt(dat = dplyr::filter(all_obs, prot == "pa"), periods = periods, res = 1000, logCount = TRUE, species = "species", x = "EASTING", y = "NORTHING", year = "year", spatialUncertainty = "spatialUncertainty", identifier = "tax_grp", output = "nPeriods", sites_gr = sites_gr, pa = TRUE)

maps_pers_tax_pa_func <- function(x, leg = FALSE) {
  
  maps_out <- maps_pers_tax_pa[[x]] +
    coord_equal() +
    # add maximum repeated value
    geom_text(x = 550000, y = 600000, label = max(maps_pers_tax_pa[[x]]$data$value, na.rm = TRUE), size = 3) +
    # change colour scheme
    ggplot2::scale_fill_manual(values = colrs_nam[2:max(maps_pers_tax_pa[[x]]$data$value, na.rm = TRUE)], na.value = "white", na.translate = FALSE, name = "Number of years sampled", drop = FALSE) +
    # edit theme
    theme(axis.ticks = element_blank(), axis.text = element_blank(), axis.title = element_blank(), panel.border = element_blank(), panel.grid = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank())

  # legend
  if(leg == FALSE) {maps_out <- maps_out + 
    theme(legend.position = "none")}
  
  return(maps_out)
   
}

maps_pers_tax_pa_out <- lapply(1:length(maps_pers_tax_pa), maps_pers_tax_pa_func)

# unprotected
maps_pers_tax_unp <- assessSpatialCov_adapt(dat = dplyr::filter(all_obs, prot == "unp"), periods = periods, res = 1000, logCount = TRUE, species = "species", x = "EASTING", y = "NORTHING", year = "year", spatialUncertainty = "spatialUncertainty", identifier = "tax_grp", output = "nPeriods", sites_gr = sites_gr, pa = FALSE)

maps_pers_tax_unp_func <- function(x, leg = FALSE) {
  
  maps_out <- maps_pers_tax_unp[[x]] +
    coord_equal() +
    # add maximum repeated value
    geom_text(x = 550000, y = 600000, label = max(maps_pers_tax_unp[[x]]$data$value, na.rm = TRUE), size = 3) +
    # change colour scheme
    ggplot2::scale_fill_manual(values = colrs_nam[2:max(maps_pers_tax_unp[[x]]$data$value, na.rm = TRUE)], na.value = "white", na.translate = FALSE, name = "Number of years sampled", drop = FALSE) +
    # edit theme
    theme(axis.ticks = element_blank(), axis.text = element_blank(), axis.title = element_blank(), panel.border = element_blank(), panel.grid = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank())
  
  # legend
  if(leg == FALSE) {maps_out <- maps_out + 
    theme(legend.position = "none")}
  
  return(maps_out)
  
}

maps_pers_tax_unp_out <- lapply(1:length(maps_pers_tax_unp), maps_pers_tax_unp_func)

# combine
maps_pers_tax_comb <- cowplot::plot_grid(cowplot::plot_grid(maps_pers_tax_pa_out[[1]], maps_pers_tax_unp_out[[1]]), cowplot::plot_grid(maps_pers_tax_pa_out[[2]], maps_pers_tax_unp_out[[2]]), cowplot::plot_grid(maps_pers_tax_pa_out[[3]], maps_pers_tax_unp_out[[3]]), cowplot::plot_grid(maps_pers_tax_pa_out[[4]], maps_pers_tax_unp_out[[4]]), cowplot::plot_grid(maps_pers_tax_pa_out[[5]], maps_pers_tax_unp_out[[5]]), cowplot::plot_grid(maps_pers_tax_pa_out[[6]], maps_pers_tax_unp_out[[6]]), ncol = 2)

# cowplot::save_plot("occ_assess/maps_pers_tax.png", maps_pers_tax_comb, base_height = 12, base_aspect_ratio = 0.7, dpi = 300)

# generate legend
maps_pers_leg <- cowplot::get_legend(maps_pers_tax_pa_func(3) + scale_fill_manual(values = colrs_nam[2:max(maps_pers_tax_pa[[3]]$data$value, na.rm = TRUE)], na.value = "white", name = "Number of\nyears sampled", drop = FALSE) + theme(legend.position = "bottom", legend.direction = "horizontal"))

# add legend to plot
maps_pers_tax_comb_leg <- cowplot::plot_grid(maps_pers_tax_comb, cowplot::plot_grid(NULL, maps_pers_leg, NULL, ncol = 3), ncol = 1, rel_heights = c(3, 0.5))

maps_pers_tax_comb_leg2 <- ggdraw(maps_pers_tax_comb_leg) +
  theme(plot.background = element_rect(fill = "white", color = NA))

# cowplot::save_plot("occ_assess/maps_pers_tax.png", maps_pers_tax_comb_leg, base_height = 12, base_aspect_ratio = 0.7, dpi = 600)

## Spatial coverage ##

# protected
maps_tax_pa <- assessSpatialCov_adapt(dat = dplyr::filter(all_obs, prot == "pa"), periods = periods_7, res = 1000, logCount = TRUE, species = "species", x = "EASTING", y = "NORTHING", year = "year", spatialUncertainty = "spatialUncertainty", identifier = "tax_grp", output = "density", sites_gr = sites_gr, pa = TRUE)

trans_col <- rgb(255,255,255, max = 255, alpha = 0)

maps_tax_pa_func <- function(x, leg = TRUE) {
  
  maps_out <- maps_tax_pa[[x]] +
    # change period names
    facet_wrap(. ~ variable, labeller = period_labeller) +
    # change fill colour
    ggplot2::scale_fill_gradient2(low = "#d9d9d9", high = "#000000", na.value = trans_col, name = expression("log"[10]*"(n records)")) +
    # edit theme
    theme(axis.ticks = element_blank(), axis.text = element_blank(), axis.title = element_blank(), panel.border = element_blank(), panel.grid = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank())
  
  # legend
  if(leg == FALSE) {maps_out <- maps_out + 
    theme(legend.position = "none")}
  
  return(maps_out)
  
}

maps_tax_pa_out <- lapply(1:length(maps_tax_pa), maps_tax_pa_func)

# unprotected
maps_tax_unp <- assessSpatialCov_adapt(dat = dplyr::filter(all_obs, prot == "unp"), periods = periods_7, res = 1000, logCount = TRUE, species = "species", x = "EASTING", y = "NORTHING", year = "year", spatialUncertainty = "spatialUncertainty", identifier = "tax_grp", output = "density", sites_gr = sites_gr, pa = FALSE)

maps_tax_unp_func <- function(x, leg = TRUE) {
  
  maps_out <- maps_tax_unp[[x]] +
    # change period names
    facet_wrap(. ~ variable, labeller = period_labeller) +
    # change fill colour
    ggplot2::scale_fill_gradient2(low = "#d9d9d9", high = "#000000", na.value = trans_col, name = expression("log"[10]*"(n records)")) +
    # edit theme
    theme(axis.ticks = element_blank(), axis.text = element_blank(), axis.title = element_blank(), panel.border = element_blank(), panel.grid = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank())
  
  # legend
  if(leg == FALSE) {maps_out <- maps_out + 
    theme(legend.position = "none")}
  
  return(maps_out)
  
}

maps_tax_unp_out <- lapply(1:length(maps_tax_unp), maps_tax_unp_func)

# combine

# ants, bees
maps_ab <- cowplot::plot_grid(maps_tax_pa_out[[1]] + labs(title = "Ants"), maps_tax_unp_out[[1]] + labs(title = "Ants"), maps_tax_pa_out[[2]] + labs(title = "Bees"), maps_tax_unp_out[[2]] + labs(title = "Bees"), ncol = 2)

cowplot::save_plot("occ_assess/maps_ab.png", maps_ab, base_height = 12, base_aspect_ratio = 0.7, dpi = 300)

# hoverflies, ladybirds
maps_hl <- cowplot::plot_grid(maps_tax_pa_out[[3]] + labs(title = "Hoverflies"), maps_tax_unp_out[[3]] + labs(title = "Hoverflies"), maps_tax_pa_out[[4]] + labs(title = "Ladybirds"), maps_tax_unp_out[[4]] + labs(title = "Ladybirds"), ncol = 2)

cowplot::save_plot("occ_assess/maps_hl.png", maps_hl, base_height = 12, base_aspect_ratio = 0.7, dpi = 300)

# spiders, wasps
maps_sw <- cowplot::plot_grid(maps_tax_pa_out[[5]] + labs(title = "Spiders"), maps_tax_unp_out[[5]] + labs(title = "Spiders"), maps_tax_pa_out[[6]] + labs(title = "Wasps"), maps_tax_unp_out[[6]] + labs(title = "Wasps"), ncol = 2)

cowplot::save_plot("occ_assess/maps_sw.png", maps_sw, base_height = 12, base_aspect_ratio = 0.7, dpi = 300)

## Spatial bias ##

spat_tax_func <- function(tax) {
  
  spatBias_tax_pa <- occAssess::assessSpatialBias(dat = dplyr::filter(all_obs, prot == "pa" & tax_grp == !!tax), periods = periods, mask = pa_rast, nSamps = 100, degrade = TRUE, species = "species", x = "EASTING", y = "NORTHING", year = "year", spatialUncertainty = "spatialUncertainty", identifier = "prot")
  
  spatBias_tax_unp <- occAssess::assessSpatialBias(dat = dplyr::filter(all_obs, prot == "unp" & tax_grp == !!tax), periods = periods, mask = pa_rast, nSamps = 100, degrade = TRUE, species = "species", x = "EASTING", y = "NORTHING", year = "year", spatialUncertainty = "spatialUncertainty", identifier = "prot")
  
  spat_dat_tax <- dplyr::bind_rows(spatBias_tax_pa$data, spatBias_tax_unp$data) %>%
    # convert period to year
    dplyr::mutate(year = as.numeric(Period) + 1989) %>% 
    dplyr::mutate(identifier = dplyr::recode(identifier, pa = "Protected", unp = "Unprotected"))
  
  spatBias_tax_out <- ggplot(data = spat_dat_tax, aes(x = year, y = mean, ymin = lower, ymax = upper, colour = identifier, fill = identifier)) + 
    geom_hline(yintercept = 1, colour = "grey", lty = 2) +
    geom_line() + 
    geom_point() + 
    geom_ribbon(alpha = 0.5) + 
    # mean value across years
    geom_text(x = 1992, y = 1.3, label = sprintf("%.2f", signif(mean(dplyr::filter(spat_dat_tax, identifier == "Protected")$mean), digits = 2)), colour = "#0077BB", size = 3) +
    geom_text(x = 1992, y = 1.1, label = sprintf("%.2f", signif(mean(dplyr::filter(spat_dat_tax, identifier == "Unprotected")$mean), digits = 2)), colour = "#EE7733", size = 3) +
    theme_linedraw() + 
    scale_colour_manual(values = c("#0077BB", "#EE7733")) +
    scale_fill_manual(values = c("#0077BB", "#EE7733")) +
    labs(title = tax, x = "Year", y = "Nearest neighbour index", colour = "", fill = "") +
    ylim(c(0, 1.4))
  
  return(spatBias_tax_out)
  
}

# runs ~1 hour
spatBias_tax <- lapply(unique(all_obs$tax_grp), spat_tax_func)

spatBias_tax_leg <- cowplot::get_legend(spatBias_tax[[1]] +theme(legend.position = "bottom"))

spatBias_tax_comb <- cowplot::plot_grid(spatBias_tax[[1]] + theme(legend.position = "none"), spatBias_tax[[2]] + theme(legend.position = "none") + labs(y = ""), spatBias_tax[[3]] + theme(legend.position = "none"), spatBias_tax[[4]] + theme(legend.position = "none") + labs(y = ""), spatBias_tax[[5]] + theme(legend.position = "none"), spatBias_tax[[6]] + theme(legend.position = "none") + labs(y = ""), ncol = 2)

spatBias_tax_comb_leg <- cowplot::plot_grid(spatBias_tax_comb, spatBias_tax_leg, ncol = 1, rel_heights = c(3, 0.3))

spatBias_tax_comb2 <- ggdraw(spatBias_tax_comb_leg) +
  theme(plot.background = element_rect(fill = "white", color = NA))

cowplot::save_plot("occ_assess/spat_bias_tax.png", spatBias_tax_comb2, base_height = 8, base_aspect_ratio = 1, dpi = 600)

## repeat visits ##

perd <- data.frame(year = 1990:2018, perd_7 = c(rep("1990:1997", 8), rep("1998:2004", 7), rep("2005:2011", 7), rep("2012:2018", 7)))

rep_vis <- all_obs %>% 
  dplyr::count(tax_grp, prot, grid, EASTING, NORTHING, year) %>% 
  # repeat visits (i.e., two visits to same site in same year)
  dplyr::filter(n >= 2) %>% 
  dplyr::left_join(perd, by = "year")

rep_plot <- function(tax, pa = NULL) {
  
  # subset data for taxonomic group
  dat <- rep_vis %>% 
    dplyr::filter(tax_grp == !!tax)
  
  if(pa == TRUE) {
    
    dat <- dplyr::filter(dat, prot == "pa")
    
  }
  
  if(pa == FALSE) {
    
    dat <- dplyr::filter(dat, prot == "unp")
    
  }
  
  # plot
  p <- ggplot2::ggplot(data = dat, aes(x = EASTING, y = NORTHING)) +
  # add GB
  geom_tile(data = sites_gr, aes(x = EASTING, y = NORTHING), fill = "grey", colour = "grey", lwd = 0, alpha = 0.4)
  
  if(pa == TRUE) {
    
    p <- p + 
    # add protected areas
  geom_tile(data = dplyr::filter(sites_gr, prot == "pa"), aes(x = EASTING, y = NORTHING), alpha = 0.5, fill = "#0077BB", colour = "#0077BB", lwd = 0)
    
  }
  
  if(pa == FALSE) {
    
    p <- p + 
      # add unprotected areas
      geom_tile(data = dplyr::filter(sites_gr, prot == "unp"), aes(x = EASTING, y = NORTHING), alpha = 0.5, fill = "#EE7733", colour = "#EE7733", lwd = 0)
    
  }
  
  p <- p +
    # grid cells with repeat visits
    geom_tile(fill = "black") +
    coord_equal() +
    labs(title = tax) +
    # split into periods
    facet_wrap(~ perd_7) +
    theme_linedraw() +
    # edit theme
    theme(axis.ticks = element_blank(), axis.text = element_blank(), axis.title = element_blank(), panel.border = element_blank(), panel.grid = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank())
  
}

rep_maps_pa <- lapply(unique(rep_vis$tax_grp), function(x) {rep_plot(tax = x, pa = TRUE)})

rep_maps_unp <- lapply(unique(rep_vis$tax_grp), function(x) {rep_plot(tax = x, pa = FALSE)})

# combine

# ants, bees
rep_ab <- cowplot::plot_grid(rep_maps_pa[[1]], rep_maps_unp[[1]], rep_maps_pa[[2]], rep_maps_unp[[2]], ncol = 2)

cowplot::save_plot("occ_assess/rep_ab.png", rep_ab, base_height = 12, base_aspect_ratio = 0.5, dpi = 300)

# hoverflies, ladybirds
rep_hl <- cowplot::plot_grid(rep_maps_pa[[3]], rep_maps_unp[[3]], rep_maps_pa[[4]], rep_maps_unp[[4]], ncol = 2)

cowplot::save_plot("occ_assess/rep_hl.png", rep_hl, base_height = 12, base_aspect_ratio = 0.5, dpi = 300)

# spiders, wasps
rep_sw <- cowplot::plot_grid(rep_maps_pa[[5]], rep_maps_unp[[5]], rep_maps_pa[[6]], rep_maps_unp[[6]], ncol = 2)

cowplot::save_plot("occ_assess/rep_sw.png", rep_sw, base_height = 12, base_aspect_ratio = 0.5, dpi = 300)

```

